\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-01-15 papers/methods/paper.tex
Updating paper

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

@@

In this paper, we lay out a new hierarchical model that extends on past work. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}. Forecasts for the impending DES three year spectroscopic supernova survey are contain in Section \ref{sec:forecast}. Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements of apparent magnitude and colour. Instead, most techniques fit these observations of magnitude (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics. As such, the product at the end is a (redshift dependent) population of $m_B$, $x_1$ and $c$.

The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology. However, given some modelled underlying population that lives in the redshift dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude. Specifically, for any given supernova our functional map may take the traditional form:
\begin{equation}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections},
\end{equation}
where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter. The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \\
\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B
\end{align}
The minimising function is then given as
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C)
\end{equation}
where $C^{-1}$ is an uncertainty matrix which combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review). The benefit this analysis methodology provides is speed - for samples of hundreds of supernova, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with two costs. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity. Secondly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data. Simulations following survey observational strategies and geometry and used to calculate the expected bias in distance modulus, which is then added onto the observational data.

@@

By building the systematic uncertainties and selection effects into the simulation package, there is vastly more freedom in how to treat and model those effects. Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms which simply cannot be expressed in a tractable likelihood. This freedom comes with a cost -- computation. The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion. For ABC methods, we must instead simulate an entire supernova population in its entirety -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts. This is an intensive process. Luckily, efficient sampling algorithms that have walkers which rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits, such as used in the BAMBIS framework {\red CITE RACHEL}.

@@

Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies Hierarchical Bayesian Models. With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model which they apply to the light curve fitting for supernova. \citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters provide a model which offers increased flexibility with reduced uncertainty over the traditional method. The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant difference in $\alpha$ and $\beta$ values from the original as well. Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017}.

@@

We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.

At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function. The difficulty arises in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.

\subsection{Observed and Latent Population}

Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values. Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$. Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.

The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$. This can be thought of as the `true' values for the supernova, and is normally distributed around the observed values such that 
\begin{equation}
P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov).
\end{equation}
As we are focused on the spectroscopically confirmed DES supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$, and the similarly that the observed mass is the true mass $m$. Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.

@@

We model the underlying population and intrinsic dispersion together as a redshift dependent multivariate skew normal for each survey. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes. Both the colour and stretch means are modelled with normal priors. The correlation between the magnitude, stretch and colour populations is represented in a correlation matrix $\rho$, which is treated with an LKJ prior. Skewness is fixed to 0 for the absolute magnitude population, but left free for colour and stretch, with normal priors centered at 0 to draw the skewness to zero if not well constrained by the data. The decision to include skewness on the stretch distribution as well as the colour was motivated by \citet{Scolnic2016}, which found highly asymmetric underlying populations for both colour and stretch, regardless of survey or scatter model adopted.

@@

As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift dependent means, width and correlations fit individual for each survey.

On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} (and therefore \citet{Kunz2007}) as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples which will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept simulation fits show that an acceptable parameterisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$. We assume that supernova fall into either population as determined by their observed classification probability $\hat{p}$.

@@

With increasingly large datasets and more nuanced analyses, the choice of how to handle $\alpha$ and $\beta$ becomes an important consideration when constructing a model. {\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0). \citet{Shariff2016} instead of employing a colour-dependent $\beta$, model $\beta$ as redshift dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.

@@

{\red Talk about how these arent even competitive with JLA and we really need to combine data. Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\usetikzlibrary{bayesnet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015}. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}. Forecasts for the impending DES three year spectroscopic supernova survey are contained in Section \ref{sec:forecasts}. {\red Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies. }

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements of apparent magnitude and colour. Instead, most techniques fit these observations of magnitude (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics. As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$.

The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology. However, given some modelled underlying population that lives in the redshift-dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude. Specifically, for any given supernova our functional map may take the traditional form:
\begin{equation}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections},
\end{equation}
where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter. The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \\
\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B
\end{align}
The minimising function is then given as
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C)
\end{equation}
where $C^{-1}$ is an uncertainty matrix that combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review). The benefit this analysis methodology provides is speed - for samples of hundreds of supernovae, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with two costs. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity. Secondly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data. As these effects are not built into the likelihood, their influence on the error budget is not captured fully in the $\chi^2$ distribution.

@@

By building the systematic uncertainties and selection effects into the simulation package, there is vastly more freedom in how to treat and model those effects. Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood. This freedom comes with a cost -- computation. The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion. For ABC methods, we must instead simulate an entire supernova population in its entirety -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts. This is an intensive process. Luckily, efficient sampling algorithms that have walkers that rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits and can help ease the computational burden.

@@

Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies Hierarchical Bayesian Models. With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model that they apply to the light curve fitting for supernova. \citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method. The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well. Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.

@@

We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift-dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.

At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.

\subsection{Observables Nature Provides: Populations}

Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values. Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$. Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$. Our set of observables is therefore given as $\lbrace \hat{m_B}, \hat{x_1}, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.

The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$. This can be thought of as the `true' values for the supernova, and is normally distributed around the observed values such that 
\begin{equation}
P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov).
\end{equation}
As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Similarly, we assume that the observed mass is the true mass $m$, such that  Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.

\begin{figure}
\begin{tikzpicture}
% Obs nodes
\node[obs]  (obssum) {$\hat{m_B}, \hat{x_1}, \hat{c}, \mathcal{C}^{-1}$};
\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};

% Latent nodes
\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
\node[det, above=of zobs] (z) {$z$};
\node[det, above=of mobs] (m) {$m$};
\node[det, above=of z] (mu) {$\mu$};
\node[latent, above=of m] (MBobs) {$M_B$};
% Survey nodes
\node[latent, above=of sum, yshift=1cm,text width=2.3cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,$\\$ \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha_c$};
\node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
% Global nodes
\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
\node[latent, above=1.7cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta,$\\$\delta(0), \delta(\infty)$};

% Connect the nodes
\edge {sum} {obssum};
\edge {z} {zobs};
\edge {m} {mobs};
\edge {z,cosmo} {mu};
\edge {mu, sum, m, pop1, ab, pobs, sf} {MBobs};

% Plates
\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(m)(mu)} {$\forall\ \rm{SN}$} ;
\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
\end{tikzpicture}
\caption{Probabilistic graphical model for our likelihood. Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.}
\label{fig:pgm}
\end{figure}

@@

We model the underlying population and intrinsic dispersion together as a redshift-dependent multivariate skew normal for each survey. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. The correlation between the magnitude, stretch and colour populations is represented in a correlation matrix $\rho$, which is treated with an LKJ prior. Skewness is fixed to 0 for the absolute magnitude and stretch, but left free for colour. We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$.

@@

As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means, width and correlations fit individually for each survey.

On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept simulation fits show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into either population as determined by their observed classification probability $\hat{p}$.

@@

With increasingly large datasets and more nuanced analyses, the choice of how to handle $\alpha$ and $\beta$ becomes an important consideration when constructing a model. {\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0). \citet{Shariff2016} instead of employ a colour-dependent $\beta$, model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.

@@

{\red Talk about how these arent even competitive with JLA and we really need to combine data? Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
